---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(LexOPS)
```
# LexOPS

<!-- badges: start -->
<!-- badges: end -->

LexOPS is an R package for generating word stimuli, for use in Psychology experiments. It can generate stimuli for a factorial design specified by the user, controlling for selected lexical variables. The package has an inbuilt database of features for English words (`LexOPS::lexops`), but the user can also use their own list of features, for English words and/or for words in other languages.

The package features an interactive shiny app, with several useful features for visualising variables' distributions and relationships while generating stimuli. The LexOPS shiny app is available online at [https://jackt.shinyapps.io/lexops/](https://jackt.shinyapps.io/lexops/), but it is usually faster and more relilable to [run it locally](#shiny-app).

## Installation

LexOPS can be installed as an R package with:

```{r, eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("JackEdTaylor/LexOPS")
```

## Shiny App

The LexOPS Shiny App can be run locally with:

```{r, eval=FALSE}
LexOPS::run_shiny()
```

## Reproducible Code

Stimuli can also be generated using reproducible code. For example, the following example pipeline generates 10 words (all nouns) per condition (60 words in total), for a study with a 2 x 3, concreteness (low, high) by emotional valence (negative, neutral, positive) experimental design. Words are controlled for by length exactly, and by word frequency within a tolerance of ±0.25 Zipf, relative to neutral abstract words.

```{r, eval=FALSE}
library(LexOPS)

stim <- lexops %>%
  subset(PoS.SUBTLEX_UK == "noun") %>%
  split_by(CNC.Brysbaert, 1:2 ~ 4:5) %>%
  split_by(VAL.Warriner, 1:3 ~ 4.5:5.5 ~ 7:9) %>%
  control_for(Zipf.SUBTLEX_UK, -0.25:0.25) %>%
  control_for(Length) %>%
  generate(n = 5, match_null = "A1_B2")

print(stim)
```

```{r, include=FALSE}
set.seed(999)
stim <- lexops %>%
  subset(PoS.SUBTLEX_UK == "noun") %>%
  split_by(CNC.Brysbaert, 1:2 ~ 4:5) %>%
  split_by(VAL.Warriner, 1:3 ~ 4.5:5.5 ~ 7:9) %>%
  control_for(Zipf.SUBTLEX_UK, -0.25:0.25) %>%
  control_for(Length) %>%
  generate(n = 5, match_null = "A1_B2")
```

```{r, echo=FALSE}
knitr::kable(stim)
```

The generated stimuli can then also be easily converted into long format, with the `long_format()` function.

```{r, eval=FALSE}
long_format(stim)
```

```{r, include=FALSE}
lf_stim <- long_format(stim)
```

```{r, echo=FALSE}
knitr::kable(lf_stim)
```
