---
title: "Higher Level Controls (control_for_map)"
author: "Jack Taylor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Higher Level Controls (control_for_map)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

The `control_for()` function works well when your variable is 1-dimensional, with a single value for each word, as it is for variables like Length, Frequency or Concreteness. Things become slightly trickier, however, when controlling for distance or similarity values, which can be calculated for each unique combination of words (i.e. $n^2$ values). One easy solution is to use `control_for_map()` to pass a function which should be used to calculate the value between any two words. Simple examples for controlling for orthographic and phonological similarity are available in the [package bookdown site](https://jackedtaylor.github.io/LexOPSdocs/advanced-stimulus-generation.html#map-functions-as-controls). This vignette demonstrates how to build your own function for `control_for_map()`, which in this example controls for semantic similarity.

# Packages

```{r, message=FALSE}
library(readr)
library(dplyr)
library(LexOPS)
```

```{r, include=FALSE}
set.seed(1)
```

# Importing Datsets

The word pair norms come from [Erin Buchanan and colleagues](https://doomlab.shinyapps.io/double_words/). This dataset indexes semantic similarity between cues and targets. Alternative sources of semantic association/relatedness values include the [Small World of Words](https://smallworldofwords.org/en/project).

```{r}
pairs <- readr::read_csv("https://media.githubusercontent.com/media/doomlab/shiny-server/master/wn_double/double_words.csv")
```

# Creating the Semantic Similarity Function

The function we create should index the similarity between $n$ matches (in a vector) and a target word (as a single string). The result should be a vector of values of length $n$, in the same order as the matches. Using the `pairs` tibble, we can use some `dplyr` manipulation to return the required values as a vector. This function returns the `root` values, indexing the cosine of two words' semantic overlap for root words (see [here](https://doomlab.shinyapps.io/double_words/) for more details).

```{r}
sem_matches <- function(matches, target) {
  # for speed, return N ones if possible
  if (all(matches==target)) return(rep(1, length(matches)))
  # find each match-target association value
  tibble(CUE = matches, TARGET = target) %>%
    full_join(pairs, by=c("CUE", "TARGET")) %>%
    filter(TARGET == target & CUE %in% matches) %>%
    pull(root)
}
```

Let's test the function on some example match-target combinations.

```{r}
# should return 1
sem_matches("yellow", "yellow")

# should return 3 values: 1 if identical, 0<x<1 if value present, NA if missing
sem_matches(c("yellow", "sun", "leaf"), "yellow")

# would return N=nrow(lexops) values (mostly NA) of similarity to "yellow"
# sem_matches(lexops$string, "yellow")
```

# Generating Stimuli

We can now generate stimuli controlling for semantic similarity. If we want to generate words which are highly semantically related, we can control for semantic relatedness of $>=0.5$ cosine similarity to an iteration's match null. Since a match null is placed at 0, and will have a similarity of 1 to itself, we set the `control_for_map()` tolerance to `-0.5:0`.

```{r}
# speed up by removing unusable word pairs from pairs
pairs <- filter(pairs, root>=0.5)

stim <- lexops %>%
  # speed up by removing strings unknown to pairs df
  dplyr::filter(string %in% pairs$CUE) %>%
  # create a random 2-level split
  split_random(2) %>%
  # control for semantic similarity
  control_for_map(sem_matches, string, -0.5:0, name = "root_cosine") %>%
  # control for other values
  control_for(Length, 0:0) %>%
  control_for(PoS.SUBTLEX_UK) %>%
  control_for(Zipf.SUBTLEX_UK, -0.2:0.2) %>%
  # generate 20 items per factorial cell (40 total)
  generate(20)
```

Here are our 20 items per factorial cell, matched by Semantic Similarity, Length, Part of Speech, and Frequency.

```{r}
print(stim)
```
