---
title: "Euclidean Distance"
author: "Jack Taylor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Euclidean Distance}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Euclidean distance can be a useful way of matching by or controlling for multiple variables with greater flexibility than dealing with the variables individually. This vignette explains how weighted and unweighted Euclidean distance is calculated in LexOPS, gives example usage with `euc_dists()` and `match_item()`, and introduces the `control_for_euc()` function. The latter allows you to generate stimuli controlling for Euclidean distance to a match null within a Generate pipeline.

# Packages

```{r packages, message=FALSE}
library(dplyr)
library(ggplot2)
library(LexOPS)
```

```{r hidden-setup, include=FALSE}
set.seed(1)
theme_set(theme_minimal())
```

# Introduction

Euclidean distance is the straight-line distance between two points in $n$-dimensional space. Between points $a$ and $b$, the Euclidean distance $d(a,b)$ is calculated as:

$$d(a,b) = \sqrt{\sum_{i=1}^n(a_i - b_i)^2}$$

This section acts as an introudction to how LexOPS functions calculate Euclidean distance, with options for scaling and weighting variables.

## An Example 2D Space

Let's imagine we want to find how close all other words are to "puppy", on two variables:

- Concreteness, from [Brysbaert et al. (2014)](http://doi.org/10.3758/s13428-013-0403-5), scaled from 1 to 5.
- Emotional valence, from [Warriner et al. (2013)](http://doi.org/10.3758/s13428-012-0314-x), scaled from 1 to 9.

The graph below represents the distance between some example words and "puppy", in this (unscaled) 2-dimensional space. The lengths of the dashed lines represent the Euclidean distance of some example words from "puppy", with the values presented in red:

```{r euc-graph1, echo = FALSE, fig.width=4, fig.height=6}
# get values for "puppy"
puppy_values <- filter(lexops, string == "puppy")
# example words of varying distances from "puppy"
example_words <- c("puppy", "trailer", "unhappiness", "cemetery", "love", "trailer", "overzealous", "juvenile", "moose")

# plot
lexops %>%
  filter(string %in% example_words) %>%
  mutate(
    cnc_dist = abs(CNC.Brysbaert - puppy_values$CNC.Brysbaert),
    val_dist = abs(VAL.Warriner - puppy_values$VAL.Warriner)
  ) %>%
  mutate(val_dist_t = 0, cnc_dist_t = 0) %>%
  bind_rows(
    mutate(., val_dist_t = val_dist, cnc_dist_t = cnc_dist)
  ) %>%
  ggplot(aes(cnc_dist, val_dist, label = string)) +
  geom_point() +
  geom_text(nudge_y = -0.25) +
  geom_line(
    aes(cnc_dist_t, val_dist_t, group = string),
    linetype = 2,
    alpha = 0.5
  ) +
  geom_text(
    aes(cnc_dist/2, val_dist/2, label = round(sqrt(cnc_dist^2 + val_dist^2), 2)),
    data = ~filter(.x, string != "puppy"),
    colour = "red"
  ) +
  coord_fixed(xlim = c(-0.1, 3.5), ylim = c(-0.1, 6.5)) +
  labs(x = "Concreteness Distance", y = "Valence Distance")
```

As mentioned, however, the above plot shows the result when our variables are not scaled. This is a problem, as the concreteness and valence variables are differently scaled. Concreteness, from [Brysbaert et al. (2014)](http://doi.org/10.3758/s13428-013-0403-5), is scaled from *1 to 5*. Emotional valence, from [Warriner et al. (2013)](http://doi.org/10.3758/s13428-012-0314-x), is scaled from *1 to 9*. If we want to give these variables equal weighting, it makes sense to scale them both first. One solution in R is the `scale()` function:

```{r scale}
lexops_scaled <- mutate(
  lexops,
  CNC.Brysbaert = scale(CNC.Brysbaert),
  VAL.Warriner = scale(VAL.Warriner)
)
```

This makes our 2-dimensional space look like the plot below. The dashed lines now reflect the values that we would get from the LexOPS function `euc_dists()`, which scales dimensions by default.

```{r euc-graph2, echo = FALSE, fig.width=6, fig.height=6}
# get values for "puppy"
puppy_values <- filter(lexops_scaled, string == "puppy")

# plot
# note that `c()` is used to remove attributes from scaled variables in puppy_values
lexops_scaled %>%
  filter(string %in% example_words) %>%
  mutate(
    cnc_dist = abs(CNC.Brysbaert - c(puppy_values$CNC.Brysbaert)),
    val_dist = abs(VAL.Warriner - c(puppy_values$VAL.Warriner))
  ) %>%
  mutate(val_dist_t = 0, cnc_dist_t = 0) %>%
  bind_rows(
    mutate(., val_dist_t = val_dist, cnc_dist_t = cnc_dist)
  ) %>%
  ggplot(aes(cnc_dist, val_dist, label = string)) +
  geom_point() +
  geom_text(nudge_y = -0.25) +
  geom_line(
    aes(cnc_dist_t, val_dist_t, group = string),
    linetype = 2,
    alpha = 0.5
  ) +
  geom_text(
    aes(cnc_dist/2, val_dist/2, label = round(sqrt(cnc_dist^2 + val_dist^2), 2)),
    data = ~filter(.x, string != "puppy"),
    colour = "red"
  ) +
  coord_fixed(xlim = c(-0.1, 4.5), ylim = c(-0.1, 4.5)) +
  labs(x = "Concreteness Distance", y = "Valence Distance")
```

To visualise in more detail what is happening, Euclidean distance in 2-dimensional space is just Pythagoras' theorem:

$$d(a,b) = \sqrt{(a_1-b_1)^2 + (a_2-b_2)^2}$$

Where $a$ is "puppy", and $b$ is "unhappiness", this can simply be represented like this:

```{r euc-graph3, echo = FALSE, fig.width=6, fig.height=6}
lexops_scaled %>%
  filter(string %in% example_words) %>%
  mutate(
    cnc_dist = abs(CNC.Brysbaert - c(puppy_values$CNC.Brysbaert)),
    val_dist = abs(VAL.Warriner - c(puppy_values$VAL.Warriner))
  ) %>%
  mutate(val_dist_t = 0, cnc_dist_t = 0) %>%
  bind_rows(
    mutate(., val_dist_t = val_dist, cnc_dist_t = cnc_dist)
  ) %>%
  filter(string %in% c("puppy", "unhappiness")) %>%
  ggplot(aes(cnc_dist, val_dist, label = string)) +
  geom_point() +
  geom_text(nudge_y = -0.1) +
  # d(a,b)
  geom_line(
    aes(cnc_dist_t, val_dist_t, group = string),
    linetype = 2,
    alpha = 0.5
  ) +
  geom_text(
    aes(cnc_dist/2, val_dist/2, label = sprintf("d(a,b) = %s", round(sqrt(cnc_dist^2 + val_dist^2), 2))),
    data = ~filter(.x, string != "puppy"),
    colour = "red"
  ) +
  # cnc
  geom_line(
    aes(cnc_dist_t, 0, group = string),
    linetype = 2,
    alpha = 0.5
  ) +
  geom_text(
    aes(cnc_dist/2, 0, label = sprintf("(a_1 - b_1) = %s", round(cnc_dist, 2))),
    data = ~filter(.x, string != "puppy"),
    colour = "red",
    nudge_y = -0.1
  ) +
  # val
  geom_line(
    aes(max(cnc_dist), val_dist_t, group = string),
    linetype = 2,
    alpha = 0.5
  ) +
  geom_text(
    aes(max(cnc_dist), val_dist/2, label = sprintf("(a_2 - b_2) = %s", round(val_dist, 2))),
    data = ~filter(.x, string != "puppy"),
    colour = "red"
  ) +
  coord_fixed(xlim = c(-0.1, 4.5), ylim = c(-0.1, 4.5)) +
  labs(x = "Concreteness Distance", y = "Valence Distance")
```

## Weights

We can also apply weights to our scaled variables, to reflect relative importance in our distance calculation. This simply consists of multiplying the scaled variables' distances by their weights $w_i$:

$$d(a,b) = \sqrt{\sum_{i=1}^n(w_i \cdot (a_i - b_i))^2}$$

As an example, if we decide to give concreteness twice the weight of valence, our 2D space would look like this:

```{r euc-graph4, echo = FALSE, fig.width=6, fig.height=6}
# get values for "puppy"
puppy_values <- lexops_scaled %>%
  mutate(CNC.Brysbaert = 2*CNC.Brysbaert) %>%
  filter(string == "puppy")

# plot
# note that `c()` is used to remove attributes from scaled variables in puppy_values
lexops_scaled %>%
  mutate(VAL.Warriner = 2*VAL.Warriner) %>%
  filter(string %in% example_words) %>%
  mutate(
    cnc_dist = abs(CNC.Brysbaert - c(puppy_values$CNC.Brysbaert)),
    val_dist = abs(VAL.Warriner - c(puppy_values$VAL.Warriner))
  ) %>%
  mutate(val_dist_t = 0, cnc_dist_t = 0) %>%
  bind_rows(
    mutate(., val_dist_t = val_dist, cnc_dist_t = cnc_dist)
  ) %>%
  ggplot(aes(cnc_dist, val_dist, label = string)) +
  geom_point() +
  geom_text(nudge_y = -0.25) +
  geom_line(
    aes(cnc_dist_t, val_dist_t, group = string),
    linetype = 2,
    alpha = 0.5
  ) +
  geom_text(
    aes(cnc_dist/2, val_dist/2, label = round(sqrt(cnc_dist^2 + val_dist^2), 2)),
    data = ~filter(.x, string != "puppy"),
    colour = "red"
  ) +
  coord_fixed(xlim = c(-0.1, 9), ylim = c(-0.1, 9)) +
  labs(x = "Concreteness Distance\n(Weight = 2)", y = "Valence Distance\n(Weight = 1)")
```

# Matching Individual Words

Imagine you want to find a close match in terms of length, frequency (Zipf), age of acquisition, and concreteness, for the word "moose". The values associated with "moose" on these variables look like this:

```{r match1, echo=FALSE}
lexops %>%
  filter(string == "moose") %>%
  select(string, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  knitr::kable()
```

You could find possible matches with variable-specific tolerances like so:

```{r match2, eval=FALSE}
lexops %>%
  match_item(
    "moose",
    Length = -1:1,
    Zipf.BNC.Written = -0.2:0.2,
    AoA.Kuperman = -2:2,
    CNC.Brysbaert = -0.25:0.25
  ) %>%
  select(string, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  head(5)
```

```{r match2-kable, echo=FALSE}
lexops %>%
  match_item(
    "moose",
    Length = -1:1,
    Zipf.BNC.Written = -0.2:0.2,
    AoA.Kuperman = -2:2,
    CNC.Brysbaert = -0.25:0.25
  ) %>%
  select(string, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  head(5) %>%
  knitr::kable()
```

Simply sorting by Euclidean distance gives a very similar result, but with some slight differences. Whereas the variable-specific tolerances excluded "hippo" (because it is >0.2 Zipf away from "moose"), in Euclidean distance it is a relatively close word and is suggested as a match.

```{r match3, eval=FALSE}
lexops %>%
  mutate(
    Euc_Dist = euc_dists(., "moose", c(Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert))
  ) %>%
  arrange(Euc_Dist) %>%
  filter(string != "moose") %>%
  select(string, Euc_Dist, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  head(5)
```

```{r match3-kable, echo=FALSE}
lexops %>%
  mutate(
    Euc_Dist = euc_dists(., "moose", c(Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert))
  ) %>%
  arrange(Euc_Dist) %>%
  filter(string != "moose") %>%
  select(string, Euc_Dist, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  head(5) %>%
  knitr::kable()
```

If we want to match by Euclidean distance in the same space, but still have strict cut-offs for frequency, we could just do the following:

```{r match4, eval=FALSE}
lexops %>%
  mutate(
    Euc_Dist = euc_dists(., "moose", c(Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert))
  ) %>%
  match_item(
    "moose",
    Euc_Dist = 0:Inf,
    Zipf.BNC.Written = -0.2:0.2
  ) %>%
  select(string, Euc_Dist, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  arrange(Euc_Dist) %>%
  head(5)
```

```{r match4-kable, echo=FALSE}
lexops %>%
  mutate(
    Euc_Dist = euc_dists(., "moose", c(Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert))
  ) %>%
  match_item(
    "moose",
    Euc_Dist = 0:Inf,
    Zipf.BNC.Written = -0.2:0.2
  ) %>%
  select(string, Euc_Dist, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  arrange(Euc_Dist) %>%
  head(5) %>%
  knitr::kable()
```

Additionally, we can set weights such that frequency accounts for more of Euclidean distance than other variables, with the `weights` argument:

```{r match5, eval=FALSE}
lexops %>%
  mutate(
    Euc_Dist = euc_dists(
      ., "moose", c(Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert),
      weights = c(1, 2, 1, 1)
    )
  ) %>%
  arrange(Euc_Dist) %>%
  select(string, Euc_Dist, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  head(5)
```

```{r match5-kable, echo=FALSE}
lexops %>%
  mutate(
    Euc_Dist = euc_dists(
      ., "moose", c(Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert),
      weights = c(1, 2, 1, 1)
    )
  ) %>%
  arrange(Euc_Dist) %>%
  select(string, Euc_Dist, Length, Zipf.BNC.Written, AoA.Kuperman, CNC.Brysbaert) %>%
  head(5) %>%
  knitr::kable()
```

# In the Generate Pipeline

Similarly, we may want to design a study comparing concrete and abstract words, controlling for length, frequency, and age of acquisition. With variable specific tolerances, our code may look like this:

```{r generate1}
stim <- lexops %>%
  split_by(CNC.Brysbaert, 1:2 ~ 4:5) %>%
  control_for(Length, 0:0) %>%
  control_for(Zipf.BNC.Written, -0.2:0.2) %>%
  control_for(AoA.Kuperman, -2:2) %>%
  generate(20)
```

This will generate stimuli with the following distributions in controls:

```{r generate1-plot, fig.width=6, fig.height=4}
plot_design(stim, "controls")
```

To generate a similar stimulus set controlling for Euclidean distance, we could use the `control_for_euc()` function. Like `euc_dists()` this has options for scaling and weighting variables. The weights here reflect the relative importance of variables as controls. Generally, variables with lower weights are permitted to vary to a greater extent.

```{r generate2}
stim2 <- lexops %>%
  split_by(CNC.Brysbaert, 1:2 ~ 4:5) %>%
  control_for_euc(
      c(Length, Zipf.BNC.Written, AoA.Kuperman),
      0:1e-5,
      name = "euclidean_distance",
      weights = c(1, 0.5, 0.1)
  ) %>%
  generate(20)
```

This will generate stimuli with the following distributions in controls. As you can see, we've generated stimuli that are closely matched in length and frequency, and more loosely matched in Age of Acquisition:

```{r generate2-plot, fig.width=6, fig.height=4}
plot_design(stim2, c("Length", "Zipf.BNC.Written", "AoA.Kuperman"))
```

The stimuli occasionally vary slightly in length in cases when age of acquisition and frequency distances are small. If we want to use this same Euclidean space, but limit how much the variables are allowed to differ, we could combine `control_for_euc()` with `control_for()`. Here we make sure that words do not differ by more than ±1 character in Length.

```{r generate3}
stim3 <- lexops %>%
  split_by(CNC.Brysbaert, 1:2 ~ 4:5) %>%
  control_for_euc(
      c(Length, Zipf.BNC.Written, AoA.Kuperman),
      0:1e-5,
      name = "euclidean_distance",
      weights = c(1, 0.5, 0.1)
  ) %>%
  control_for(Length, -1:1) %>%
  generate(20)
```

Which will give us the following distributions:

```{r generate3-plot, fig.width=6, fig.height=4}
plot_design(stim3, c("Length", "Zipf.BNC.Written", "AoA.Kuperman"))
```

We could use a similar method to give variable-specific tolerances for Zipf or Age of Acquisition if we wanted, or any variables we want, regardless of whether they're in our Euclidean space.

# Deciding on a Tolerance

Something that is relatively difficult is deciding on a tolerance for Euclidean distance. As well as just playing around with different values, it is often useful to see a sample distribution possible Euclidean distance values by generating a large number of stimuli, with no upper bound in the tolerance. Viewing the distribution can act as a first pass in terms of deciding where to set a cut-off.

```{r tolerance-space-plot, message=FALSE, fig.width=6, fig.height=4}
lexops %>%
  split_by(CNC.Brysbaert, 1:2 ~ 4:5) %>%
  control_for_euc(
      c(Length, Zipf.BNC.Written, AoA.Kuperman),
      0:Inf,
      name = "euclidean_distance",
      weights = c(1, 0.5, 0.1)
  ) %>%
  generate(500) %>%
  long_format() %>%
  ggplot(aes(as.numeric(euclidean_distance), fill = condition)) +
  geom_density(alpha = 0.5)
```

One last thing to note is that generating stimuli controlled for Euclidean distance can make your stimulus generation more flexible, but also much slower (as you can see from the console output for *total iterations* and *success rate*). This is because LexOPS cannot use some of the heuristics it applies when matching items with variable-specific tolerances to exclude inappropriate matches, and because Euclidean distance has to be recalculated each iteration (in fact, `control_for_euc()` is just a wrapper for `control_for_map()`). As a result, variable-specific tolerances may be more useful if computational efficiency is important.
